---
title: "03_preprocess_data"
author: "Jae Yeon Kim"
date: "7/2/2020"
output: html_document
---

# Import packages and files 

## Packages

```{r}

pacman::p_load(data.table, # for fast data manipulation
               tidyverse, # for tidyverse 
               ggpubr, # for arranging ggplots   
               ggthemes, # for fancy ggplot themes
               here, # for reproducibility 
               countrycode, # to convert country code <-> country name 
               maps, # for US city data
               openintro, # for revert state name abbreviations to their original forms 
               stringr, # for easy regular expression
               lubridate, # for easy time var manipulation
               forcats, # reverse factor order 
               patchwork, # for easy ggarrange
               ggsci) # for pubs 

devtools::install_github("jaeyk/tidytweetjson", force = TRUE)

library(tidytweetjson)

# for publication-friendly theme 
theme_set(theme_pubr())

# custom functions
source(here("functions", "stacked_bar_plot.R"))

```

## Files 

- 5,050,042 obs 

```{r}

parsed <- readRDS(here("processed_data", "parsed.rds"))

```

# Wrangle 

## Filter

- Filtered if location field is empty or NA: 3,734,756 obs (73% of the original data)
- I did not use country_code field as it is mostly filled with NAs (97%).

```{r}


sum(is.na(parsed$country_code))/nrow(parsed)

filtered <- parsed %>%
    filter(!(location == "" | is.na(location)))

```

## Select 

```{r}

# Select columns 
selected <- filtered %>% 
    select(country_code, location, created_at, full_text) 

```

## Mutate 

### Location 

- `maps::us.cities` database includes the names of the US cities whose population sizes are larger than 40,000. 

- 41% of the Tweets were created by the users located in the US.

```{r}

# Identify US location 

us_cities <- maps::us.cities$name %>% str_replace_all(" [[:alpha:]]*$", "") %>% unique() %>% tolower()

us_states <- maps::us.cities$name %>% stringr::word(-1) %>% unique() %>% abbr2state() %>% tolower()

us_country <- c("United States", "USA", "US", "U.S.A.") %>% tolower()

us_location_list <- c(us_cities, us_states, us_country)

```

```{r}

selected$US <- str_detect(selected$location %>% tolower(), us_location_list %>% paste(collapse = "|")) %>% as.numeric()

sum(selected$US)/nrow(selected)

```

```{r}

df <- selected %>% filter(US == 1)

fwrite(df, here("processed_data", "US_tweets.csv"))

df <- fread(here("processed_data", "US_tweets.csv"))

```

### Text 

```{r}

# Create new columns 
df <- df %>%
    mutate(full_text = str_to_lower(full_text),
           wuhan = as.numeric(str_detect(full_text, "wuhan")),
           asian = as.numeric(str_detect(full_text, "asia|asian")),
           chinese = as.numeric(str_detect(full_text, "china|chinese")))

```

### Date

```{r}
df <- tidytweetjson::add_date(df)

#df[1:10,] %>% select(created_at, date)
```

# Explore 

## Time-series

### Overall 

```{r}

# Group by date then visualize 
df %>%
    group_by(date) %>%
    count() %>%
    ggplot(aes(x = as.Date(date), y = log(n), group = 1)) +
        geom_line() +
        labs(x = "Date",
             y = "Logged count",
             title = "The Count of Tweets on COVID-19",
             subtitle = "The Tweets were created by the users located in the US")

ggsave(here("outputs", "overall_trend.png"), height = 7)

```

### Keywords 

```{r}

p1 <- stacked_area_plot(df, wuhan) +
        labs(x = "Date",
             y = "Proportion",
             fill = "Wuhan",
             title = "Wuhan",
             subtitle = "The Tweets were created by the users located in the US")

p2 <- stacked_area_plot(df, chinese) +
        labs(x = "Date",
             y = "Proportion",
             fill = "Chinese",
             title = "Chinese",
             subtitle = "The Tweets were created by the users located in the US")

p3 <- stacked_area_plot(df, asian) +
        labs(x = "Date",
             y = "Proportion",
             fill = "Asian",
             title = "Asian",
             subtitle = "The Tweets were created by the users located in the US")

p1 / p2 / p3

ggsave(here("outputs", "stacked_bar_plots.png"),
       height = 10)
```

# Export 

```{r}

fwrite(df, here("processed_data", "text_ready.csv"))

```

